Measurement of total hadronic differential cross sections in the LArIAT experiment

### 2018-04-19 Chapter-2/2-LArTPCs.tex
s

\subsection{Time Projection Chamber}
\subsection{Ionization Detectors with Noble Liquids}
\subsection{LArTPC: Principles of Operation}\label{sec:LArTPCWorkingPrinciple}
Details on the production and detection of ionization charge and scintillation light are provided in \ref{sec:light} and \ref{sec:light} respectively.
\subsection{Liquid Argon Ionization Charge Detection}\label{sec:charge}
\subsubsection{Electron Life Time \& purity}
LArTPCs use  hermetically sealed and leak-checked vessels to abate the leakage and diffusion of contaminants into the system. The liquid argon filling of the volume occurs after the vessel is evacuated or purged with gaseous argon [205]  to reduce remaining gases in the volume. Even so, the construction of a pure tank of argon is unviable, as several sources of impurity remain.  In particular, impurities can come from the raw argon supply, the argon filtration system and from the outgassing from internal surfaces. Outgassing is a continuous diffusive process  producing contaminants, especially water, even after the vessel is sealed, particularly from materials in the ullage region\footnote{.  While the liquid argon low temperature reduces outgassing in the liquid, this process remains significant for absorptive material (such as plastic) above the surface of the liquid phase}.  Since research-grade argon comes from the industrial distillation of air, the impurities with the highest concentration are nitrogen, oxygen and water, generally maintained under the 1 part per million level by the vendor.  Even so, a higher level of purity is necessary to achieve a free electron life time usable in meter scale detectors. Thus, argon  is constantly  filtered in the cryogenic system, which reduce the oxygen and water contamination to less than 100 parts per trillion. The filtration system depends on the size and drift distance of the experiment and, for experiments on several meters scale, it includes an argon recirculation system.

\subsubsection{Space Charge Effect}
\subsubsection{Recombination Effect}
\subsection{Liquid Argon Scintillation Light }\label{sec:light}
Liquid argon emits scintillation light at the passage of charged particles. LArTPCs  leverage this property to determine when the ionization charge begins to drift towards the anode plane. %Scintillation light can also be used for particle identification, as discussed in the next sections.

@@

The stringent constraints for the electron life time limit the presence of oxygen and water to such a low level that both absorption and quenching on these impurity is not expected to be significant [210].   Contrarily, the nitrogen level is not bound by the electron life time constraints -- nitrogen being an inert gas, expensive to filter. Thus, nitrogen is often present at the level provided by the vendor. For a nitrogen  concentration of 2 parts per million,  typical of the current generation of LArTPC, the attenuation length due to nitrogen has been measured to be $\sim$30 meters \cite{1748-0221-8-07-P07011}.

\textcolor{red}{ADD quenching effects}

\subsubsection{Wavelength Shifting of LAr Scintillation Light}
\subsection{Signal processing}
\section{The SBN Program: Neutrino Interaction and Detection}
%\subsection{SBN Goals}
%\subsection{Neutrino Interactions and Detection }
\section{DUNE: Rare Decay Searches}
The key elements for a rare decay experiment are: massive active volume, long exposure, high identification efficiency and low background. 
%The limit to proton lifetime in case of absence of signal and backgrounds is set by calculating
%$$\tau/B > M\times \epsilon\times T \times 10^{32},$$ 
%where M is the detector mass in kton, $\epsilon$ the signal detection efficiency after cuts to suppress backgrounds (dependent on the considered decay mode), T is the exposure in years, B the assumed branching fraction for the considered mode and  $10^{32}$ is a factor accounting for the number of nucleons in a kton of material \cite{Bueno2007}.
Figure \ref{fig:PDKExperimentalLImit} shows the current best experimental limits on nucleon decay lifetime over branching ratio (dots). Historically, the dominant technology used in these searches has been water Cherenkov detectors: all the best experimental limits on every decay mode are indeed set by Super-Kamiokande \cite{PhysRevD.90.072005,PhysRevLett.115.121803}.  It is particularly important to notice that the kaon energy for the proton decay mode $p \rightarrow K^+ \bar{\nu}$ is under Cherenkov threshold.  Super-Kamiokande set the limit on the lifetime for the $p \rightarrow K^+ \bar{\nu}$ mode by  relying exclusively on photons from nuclear de-excitation. For this reason, an attractive alternative approach to identifying nucleon decay is the use of a Liquid Argon Time Projection Chamber (LArTPC).

>>>>

\subsection{Time Projection Chamber}
\subsection{Ionization Detectors with Noble Liquids}
\subsection{LArTPC: Principles of Operation}\label{sec:LArTPCWorkingPrinciple}

%The time projection chamber (TPC) was invented by Nygren in 1974 [91]. The proposal to implement a liquid argon TPC (LArTPC) for neutrino physics was made by Rubbia in 1977 [92], and the concept was implemented by the ICARUS collaboration [93].

To the bare bones, a LArTPC is a bulk of liquid argon sandwiched in a flat capacitor, equipped with a light collection system. A uniform electric field of the order of ~500 V/cm is maintained constant between the faces of the capacitor.
Two processes occurs when an ionizing radiation traverses the argon active volume: the argon ionize and scintillates along the charge particle's trajectory. Details on the production and detection of ionization charge and scintillation light are provided in \ref{sec:light} and \ref{sec:light} respectively. The optical part of the detector collects the argon scintillation light in matters of nanoseconds. This flash of light flags the starting time of an event in the chamber, $t_0$. The uniform electric field drifts the electrons from ionization towards the anode plane and the ionized argon towards the cathode.

\subsection{Liquid Argon Ionization Charge Detection}\label{sec:charge}
\subsubsection{Electron Life Time \& purity}
LArTPCs use  hermetically sealed and leak-checked vessels to abate the leakage and diffusion of contaminants into the system. The liquid argon filling of the volume occurs after the vessel is evacuated or purged with gaseous argon [205]  to reduce remaining gases in the volume. Even so, the construction of a pure tank of argon is unviable, as several sources of impurity remain.  In particular, impurities can come from the raw argon supply, the argon filtration system and from the outgassing from internal surfaces. Outgassing is a continuous diffusive process  producing contaminants, especially water, even after the vessel is sealed, particularly from materials in the ullage region\footnote{.  While the liquid argon low temperature reduces outgassing in the liquid, this process remains significant for absorptive material (such as plastic) above the surface of the liquid phase}.  Since research-grade argon comes from the industrial distillation of air, the impurities with the highest concentration are nitrogen, oxygen and water, generally maintained under the 1 part per million level by the vendor.  Even so, a higher level of purity is necessary to achieve a free electron life time usable in meter scale detectors. Thus, argon  is constantly  filtered in the cryogenic system, which reduce the oxygen and water contamination to less than 100 parts per trillion. The filtration system depends on the size and drift distance of the experiment and, for experiments on several meters scale, it includes an argon recirculation system.

\subsubsection{Space Charge Effect}
\subsubsection{Recombination Effect}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%                     Light Detection             %%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Liquid Argon Scintillation Light }\label{sec:light}
Liquid argon emits scintillation light at the passage of charged particles. LArTPCs  leverage this property to determine when the ionization charge begins to drift towards the anode plane. %Scintillation light can also be used for particle identification, as discussed in the next sections.

@@

The stringent constraints for the electron life time limit the presence of oxygen and water to such a low level that both absorption and quenching on these impurity is not expected to be significant [210].   Contrarily, the nitrogen level is not bound by the electron life time constraints -- nitrogen being an inert gas, expensive to filter. Thus, nitrogen is often present at the level provided by the vendor. The effects of nitrogen on argon scintillation light have been studied in the WArP R\&D program and at several test stands.
The quenching process induced by nitrogen in liquid Ar has been measured to be proportional to the nitrogen concentration, with a rate constant of $\sim  0.11$ $\mu s^{-1}$ ppm$^{-1}$; appreciable decreasing in lifetime and relative amplitude of the slow component have been shown for contamination as high as a few ppm of nitrogen \cite{1748-0221-5-06-P06003}.
For a nitrogen  concentration of 2 parts per million,  typical of the current generation of LArTPC, the attenuation length due to nitrogen has been measured to be $\sim$30 meters \cite{1748-0221-8-07-P07011}.

\subsubsection{Wavelength Shifting of LAr Scintillation Light}
Liquid argon scintillation light is invisible for most optical detectors deployed in LArTPC, such as cryogenic PMTs and SiPMs, since a wavelength of 128 nm is  generally too short to be absorbed from most in glasses, polymers and semiconductor materials. Research on prototype SiPMs absorbing directly VUV light and their deployment in noble gasses experiment is ongoing but not mature \cite{1748-0221-8-01-C01003}. Thus, experiments need to shift the wavelength of scintillation light to be able to detect it.  Albeit deployed in different ways, neutrinos and dark matter experiments commonly use  1,1,4,4-tetraphenyl-butadiene (TPB) to shift the scintillation light. 
TPB, whose chemical structure is shown in figure , absorbs the vacuum ultraviolet (VUV) light and emits in the visible at $\sim$~425 nm [231], with a ratio of visible photon emitted per VUV photon absorbed of $\sim$1.2:1 \cite{GEHMAN2011116}.

Neutrino experiments typically coat their optical detector system evaporating a layer of TPB either directly on the PMTs glass surface or on acrylic plates mounted in front of the PMTs \textcolor{red}{cite microboone}; this technique allows the fast detection light coming directly from the neutrino interaction. Dark matter experiments typically evaporate TPB on reflective foils mounted on the inside walls of the sensitive volume and detect the light after it has been reflected; this technique leads to a higher and more uniform light yield, though scattering effects for both the visible and VUV light augment the propagation time and hinder directionality information\textcolor{red}{some DM detector} . In order to take advantage of both these techniques, hybrid systems with PMT coating and foils are being considered for the next generation of large neutrino detectors \textcolor{red}{cite SBND?}.

\subsection{Signal processing}
\section{The SBN Program: Neutrino Interaction and Detection}
%\subsection{SBN Goals}
%\subsection{Neutrino Interactions and Detection }
\section{DUNE: Rare Decay Searches}
The key elements for a rare decay experiment are: massive active volume, long exposure, high identification efficiency and low background. 
%The limit to proton lifetime in case of absence of signal and backgrounds is set by calculating
%$$\tau/B > M\times \epsilon\times T \times 10^{32},$$ 
%where M is the detector mass in kton, $\epsilon$ the signal detection efficiency after cuts to suppress backgrounds (dependent on the considered decay mode), T is the exposure in years, B the assumed branching fraction for the considered mode and  $10^{32}$ is a factor accounting for the number of nucleons in a kton of material \cite{Bueno2007}.
Figure \ref{fig:PDKExperimentalLImit} shows the current best experimental limits on nucleon decay lifetime over branching ratio (dots). Historically, the dominant technology used in these searches has been water Cherenkov detectors: all the best experimental limits on every decay mode are indeed set by Super-Kamiokande \cite{PhysRevD.90.072005,PhysRevLett.115.121803}.  It is particularly important to notice that the kaon energy for the proton decay mode $p \rightarrow K^+ \bar{\nu}$ is under Cherenkov threshold.  Super-Kamiokande set the limit on the lifetime for the $p \rightarrow K^+ \bar{\nu}$ mode by  relying exclusively on photons from nuclear de-excitation. For this reason, an attractive alternative approach to identifying nucleon decay is the use of a Liquid Argon Time Projection Chamber (LArTPC).
